---
title: "Data cleaning"
output:
  pdf_document: default
  html_document: default
date: "2026-02-23"
---


## Data cleaning for initial comparison of reproductive output between sites.
This document outlines how the raw data was processed. The output is two plots. The first shows a comparison of sandprawn (Kraussillichirus kraussi) egg biomass across the five sites that were sampled. The second is a comparison of sandprawn embryo numbers across the sites.

### Reading in the data

```{r results='hide'}
library(readxl)
rawdata <- read_excel("Ella sandprawn data.xlsx")
head(rawdata)
tail(rawdata)
```

### Selecting the relevant coloumns: Site, Number of embryos and Egg biomass

```{r}
spdata <- rawdata[, c(3, 10, 11)]
head(spdata)
str(spdata)
```

### Converting Site from numeric to a factor

``` {r}
library(tidyverse)
spdata <- spdata %>% mutate(Site= as.factor(Site))
str(spdata)
```

### Removing observations where animals did not have eggs

``` {r}
clean_spdata <- spdata %>% filter(`Egg biomass (g)` > 0)
head(clean_spdata)
summary(clean_spdata)
``` 

### Creating boxplot of sandprawn egg biomass across sites

``` {r}
library(ggplot2)
ggplot(data = clean_spdata, aes(x = Site, y = `Egg biomass (g)`)) + geom_boxplot() + labs(title = "Comparison of Sandprawn egg biomass across sites", x = "Site collected", y = " Egg biomass (g)")
```

### Creating boxplot of sandprawn embryo numbers across sites

``` {r}
ggplot(data = clean_spdata, aes(x = Site, y = `No. of embryos`)) + geom_boxplot() + labs(title = "Comparison of Sandprawn embryo numbers across sites", x = "Site collected", y = " Number of embryos")
```

